---
title: "Chapter 7 (III) - Sensitivity Analysis"
output: html_notebook
author: Torran Semple
date: 06/2025
---

## Load synth data (version including fuel poverty estimates data)
```{r}

# Load data
synth_fp <- read.csv("/Users/psxts6/Desktop/FP_income_PoC/code_and_data/synth_plus_fp_estimates.csv", header = TRUE, sep = ",")

# This file contains synthetic Nottingham homes and FP estimates resulting from CH7; the following chunks prepare the data for sensitivity analysis 1 and 2

```


## Sensitivity analysis 1 (SA1): potential energy efficiency
```{r}

# Notes: SA1 assumes that all homes reach their "potential energy efficiency" and that income and energy unit prices are held constant; this requires the estimation of "potential energy costs" based on changing consumption variables, following a similar process to that described to infer energy costs in the Chapter 7 analysis, as shown below:


# Step i) total potential consumption
synth_fp$total_consumption_potential <- synth_fp$energy_consumption_potential * synth_fp$total_floor_area


# Step ii) energy split 
gas_factor <- 0.8099
electricity_factor <- 0.1901

synth_fp$gas_consumption_potential <- synth_fp$total_consumption_potential * gas_factor 
synth_fp$electricity_consumption_potential <- synth_fp$total_consumption_potential * electricity_factor


# Step iii) annual gas and electricity costs based on Ofgem price caps & standing charge
gas_unit <- 0.0634
electricity_unit <- 0.2486
gas_daily <- 0.3165
electricity_daily <- 0.6097

synth_fp$gas_costs_potential <- (synth_fp$gas_consumption_potential * gas_unit) + (365 * gas_daily) 
synth_fp$electricity_costs_potential <- (synth_fp$electricity_consumption_potential * electricity_unit) + (365 * electricity_daily)


# Step iv) total potential costs = energy_costs_SA1: 
synth_fp$energy_costs_SA1 <- synth_fp$gas_costs_potential + synth_fp$electricity_costs_potential


# compare potential energy costs (SA1) to current energy costs
summary(synth_fp$energy_costs)
summary(synth_fp$energy_costs_SA1)



# Note relevant variables for SA1 are: energy_costs_SA1; income (unchanged); and potential_energy_efficiency (i.e., improved SAP scores) 


```


## Sensitivity analysis 2: worst case scenario (energy costs rise faster than income)
```{r}

# Create new columns for energy_costs_SA2 (+78.38%) and income_SA2 (+1.46%)
synth_fp$income_SA2 <- synth_fp$income * 1.0146
synth_fp$energy_costs_SA2 <- synth_fp$energy_costs * 1.7838

```


## Complete summary of SA parameters
```{r}

# Reference (synthetic Nottingham homes)
summary(synth_fp$income)
summary(synth_fp$energy_costs)
summary(synth_fp$current_energy_efficiency)


# SA1
summary(synth_fp$income)
summary(synth_fp$energy_costs_SA1)
summary(synth_fp$potential_energy_efficiency)


# SA2
summary(synth_fp$income_SA2)
summary(synth_fp$energy_costs_SA2)
summary(synth_fp$current_energy_efficiency)


```


## Sensitivity analysis 1 (SA1): All homes reach "potenital" energy efficiency; energy unit prices and income held constant
```{r}

# Recalculate 10%, LIHC and LILEE using SA1 variables

# Calculate SA1 version of 10% (based on FP ratio for SA1 conditions; note: 20% fuel poverty also included, as may be relevant to later stages)
library(dplyr)
synth_fp <- synth_fp %>%
    mutate(fuel_poverty_ratio_SA1 = energy_costs_SA1 / income * 100) # note income variable remains unchanged; hence, original "income" used

# new columns: categorical 10% and 20%
synth_fp <- synth_fp %>%
  mutate(categorical_10_SA1 = ifelse(fuel_poverty_ratio_SA1 >= 10.0, 1, 0))  
synth_fp <- synth_fp %>%
  mutate(categorical_20_SA1 = ifelse(fuel_poverty_ratio_SA1 >= 20.0, 1, 0))  

table(synth_fp$categorical_10_SA1) # 17092 homes (16.19%) 10% fuel poor given SA1 conditions; this is considerably less than baseline 10% fuel poverty (36.83%)



# Calculate SA1 version of LIHC (values of poverty line (18840) and median energy costs across UK (2252) assumed to remain constant - i.e., the same as CH7 estimations)
synth_fp <- synth_fp %>%
  mutate(LIHC_SA1 = ifelse(energy_costs_SA1 > 2252 & income < 18840 + energy_costs_SA1, 1, 0))

table(synth_fp$LIHC_SA1) # As expected, rate of LIHC fuel poverty is much lower in SA1 (n=2652, 2.51%) than baseline LIHC (10.26%)




# Calculate the SA1 version of LILEE* - pov_line_2021 still = 18840 as above for LIHC, and use "potential" energy efficiency

# poverty line 2021 + energy costs
synth_fp <- synth_fp %>%
  mutate(pov_line_2021_energy_costs_SA1 = pov_line_2021 + energy_costs_SA1)

# LILEE* classification
synth_fp <- synth_fp %>%
  mutate(LILEE_SA1 = ifelse(potential_energy_efficiency < 69 & income < pov_line_2021_energy_costs_SA1, 1, 0))

table(synth_fp$LILEE_SA1) # As expected, rate of LILEE* is much lower in SA1 (n=3949, 3.74%) than baseline LILEE* (18.30%) 




# Calculate the SA1 version of 10% + MIS
synth_fp <- synth_fp %>%
  mutate(mis_ninety = mis_budget * 0.9)


# MIS-based fuel poverty
synth_fp <- synth_fp %>%
  mutate(
    mis_fuel_poverty_SA1 = ifelse((income - energy_costs_SA1) < mis_ninety, 1, 0)
  )


# MIS + 10% (Scot Gov. approach)
synth_fp <- synth_fp %>%
  mutate(
    mis_plus_10_SA1 = case_when(
      mis_fuel_poverty_SA1 == 1 | categorical_10_SA1 == 1 ~ 1,
      TRUE ~ 0
    )
  )

table(synth_fp$mis_plus_10_SA1) # 10% + MIS for SA1 = 24.26%; ~40% reduction compared to baseline (40.41%)


```


## Sensitivity analysis 2 (SA2): No change to housing stock energy efficiency; energy unit price appreciation outgrows income (as specified previously, energy_costs_SA2 (+78.38%) and income_SA2 (+1.46%) vs baseline conditions)
```{r}

# Recalculate 10%, LIHC and LILEE using SA2 variables

# Calculate SA2 version of 10% (based on FP ratio for SA2 conditions; note: 20% fuel poverty also included, as may be relevant to later stages)
synth_fp <- synth_fp %>%
    mutate(fuel_poverty_ratio_SA2 = energy_costs_SA2 / income_SA2 * 100) # note income variable remains unchanged; hence, original "income" used

# new columns: categorical 10% and 20%
synth_fp <- synth_fp %>%
  mutate(categorical_10_SA2 = ifelse(fuel_poverty_ratio_SA2 >= 10.0, 1, 0))  
synth_fp <- synth_fp %>%
  mutate(categorical_20_SA2 = ifelse(fuel_poverty_ratio_SA2 >= 20.0, 1, 0))  

table(synth_fp$categorical_10_SA2) # 82882 homes (78.51%) 10% fuel poor given SA2 conditions; this is considerably higher than baseline 10% fuel poverty (36.83%)



# Calculate SA2 version of LIHC (values of poverty line (18840) and median energy costs across UK (2252) assumed to remain constant - i.e., the same as CH7 estimations)
synth_fp <- synth_fp %>%
  mutate(LIHC_SA2 = ifelse(energy_costs_SA2 > 2252 & income_SA2 < 18840 + energy_costs_SA2, 1, 0))

table(synth_fp$LIHC_SA2) # Rate of LIHC fuel poverty is relatively high in SA2 (n=35212, 33.35%); however, this is partly due to the assumption of a stable UK-wide 

# NOTE: Demonstration of ALT LIHC SA2 estimate, where median energy costs (for the UK) are assumed to be £4,092 (see Chapter 8: Methods)
synth_fp <- synth_fp %>%
  mutate(LIHC_SA2_WCS = ifelse(energy_costs_SA2 > 4092 & income_SA2 < 18840 + energy_costs_SA2, 1, 0))

table(synth_fp$LIHC_SA2_WCS) # Now LIHC remains more stable (12.08%)



# Calculate the SA2 version of LILEE* - pov_line_2021 still = 18840 as above for LIHC, and use "potential" energy efficiency

# poverty line 2021 + energy costs
synth_fp <- synth_fp %>%
  mutate(pov_line_2021_energy_costs_SA2 = pov_line_2021 + energy_costs_SA2)

# LILEE* classification
synth_fp <- synth_fp %>%
  mutate(LILEE_SA2 = ifelse(current_energy_efficiency < 69 & income_SA2 < pov_line_2021_energy_costs_SA2, 1, 0))

table(synth_fp$LILEE_SA2) # Rate of LILEE* in SA2 (21.41%%) is only marginally higher than baseline LILEE* (18.30%) 




# Calculate the SA2 version of 10% + MIS
synth_fp <- synth_fp %>%
  mutate(mis_ninety = mis_budget * 0.9)


# MIS-based fuel poverty
synth_fp <- synth_fp %>%
  mutate(
    mis_fuel_poverty_SA2 = ifelse((income_SA2 - energy_costs_SA2) < mis_ninety, 1, 0)
  )


# MIS + 10% (Scot Gov. approach)
synth_fp <- synth_fp %>%
  mutate(
    mis_plus_10_SA2 = case_when(
      mis_fuel_poverty_SA2 == 1 | categorical_10_SA2 == 1 ~ 1,
      TRUE ~ 0
    )
  )

table(synth_fp$mis_plus_10_SA2) # 10% + MIS for SA2 = 78.75%; around twice the baseline rate (40.41%)



```


## Generate LSOA-level summaries for SA1 and SA2 fuel poverty estimates
```{r}

# SA1 summaries

# SA1 10% summary
summary_ten_SA1 <- synth_fp %>%
  group_by(lsoa_2011) %>%
  summarise(
   proportion_10_SA1 = mean(categorical_10_SA1, na.rm = TRUE)
  )

# SA1 LIHC summary
summary_lihc_SA1 <- synth_fp %>%
  group_by(lsoa_2011) %>%
  summarise(
   proportion_lihc_SA1 = mean(LIHC_SA1, na.rm = TRUE)
  )

# SA1 LILEE summary
summary_lilee_SA1 <- synth_fp %>%
  group_by(lsoa_2011) %>%
  summarise(
   proportion_lilee_SA1 = mean(LILEE_SA1, na.rm = TRUE)
  )

# SA1 10 + MIS summary
summary_ten_mis_SA1 <- synth_fp %>%
  group_by(lsoa_2011) %>%
  summarise(
   proportion_mis_plus_10_SA1 = mean(mis_plus_10_SA1, na.rm = TRUE)
  )




# SA2 summaries

# SA2 10% summary
summary_ten_SA2 <- synth_fp %>%
  group_by(lsoa_2011) %>%
  summarise(
   proportion_10_SA2 = mean(categorical_10_SA2, na.rm = TRUE)
  )

# SA2 LIHC summary
summary_lihc_SA2 <- synth_fp %>%
  group_by(lsoa_2011) %>%
  summarise(
   proportion_lihc_SA2 = mean(LIHC_SA2, na.rm = TRUE)
  )

# SA2 LILEE summary
summary_lilee_SA2 <- synth_fp %>%
  group_by(lsoa_2011) %>%
  summarise(
   proportion_lilee_SA2 = mean(LILEE_SA2, na.rm = TRUE)
  )

# SA2 10 + MIS summary
summary_ten_mis_SA2 <- synth_fp %>%
  group_by(lsoa_2011) %>%
  summarise(
   proportion_mis_plus_10_SA2 = mean(mis_plus_10_SA2, na.rm = TRUE)
  )



# Merge as one table
data_frames_list <- list(summary_ten_SA1, summary_lihc_SA1, summary_lilee_SA1, summary_ten_mis_SA1, summary_ten_SA2, summary_lihc_SA2, summary_lilee_SA2, summary_ten_mis_SA2)

lsoa_SA1_SA2 <- Reduce(function(x, y) full_join(x, y, by = "lsoa_2011"), data_frames_list)


```


## Join LSOA-level SA1 and SA2 estimates with existing LSOA summary
```{r}

# Load LSOA summary
lsoa_summary <- read.csv("/Users/psxts6/Desktop/FP_income_PoC/code_and_data/synth_fp_experiments_lsoa_summaries.csv", header = TRUE, sep = ",")

colnames(lsoa_summary)[1] <- "lsoa_2011"
colnames(lsoa_summary)[2] <- "proportion_10"

# Retain basic columns initially (i.e., 10%, LIHC, LILEE, 10% + MIS) and append "baseline" for comparison with SA1 and SA2 values
lsoa_summary_filtered <- lsoa_summary %>%
  select(lsoa_2011,
         proportion_10_baseline = proportion_10,
         proportion_lihc_baseline = proportion_lihc,
         proportion_lilee_baseline = proportion_lilee,
         proportion_mis_plus_10_baseline = proportion_mis_plus_10)


# Join baseline, SA1 and SA2 rates
lsoa_base_SA1_SA2 <- lsoa_summary_filtered %>%
  left_join(lsoa_SA1_SA2, by = "lsoa_2011")


# write.csv(x=lsoa_base_SA1_SA2, file="lsoa_summary_base_SA1_SA2.csv")


```


## Visualisations for baseline, SA1 and SA2 rates of fuel poverty
```{r}

lsoa_summary <- lsoa_base_SA1_SA2

# Multiply all numerical columns by 100 (i.e resolve props to percentages)
lsoa_summary <- lsoa_summary %>% mutate(across(2:13, ~ .x * 100))

# Load necessary libraries
library(dplyr)
library(tidyr)
library(ggplot2)
library(cowplot)

# 10% colour = "#DDB0D9"
# 10% + MIS colour = "#BB64B1"
# LIHC colour = "#ACE4E4"
# LILEE colour = "#5A97BC"


## --- Plot 1: proportion_10_baseline vs. proportion_10_SA1 ---
plot1 <- ggplot(data = lsoa_summary, aes(x = proportion_10_baseline, y = proportion_10_SA1)) +
  geom_point(alpha = 0.9, color = "#DDB0D9") +
  geom_point(size = 3, alpha = 0.9, shape = 21, stroke = 0.25, color = "black") +
  geom_abline(intercept = 0, slope = 1, linetype = "dashed", color = "black") +
  scale_x_continuous(breaks = seq(0, 100, 10), limits = c(0, 100)) +
  scale_y_continuous(breaks = seq(0, 100, 10), limits = c(0, 100)) +
  labs(
    title = "LSOA-level 10% fuel poverty: Baseline vs. SA1 conditions",
    x = "Baseline fuel poverty (%)",
    y = "SA1 fuel poverty (%)"
  ) +
  theme_bw() +
  theme(axis.text.x = element_text(size = 17),
        axis.text.y = element_text(size = 17),
        legend.text = element_text(size = 17),
        legend.title = element_text(size = 17),
        plot.title = element_text(size = 18),
        axis.title.x = element_text(size = 17),
        axis.title.y = element_text(size = 17)) +
  coord_fixed()

# print(plot1)

# --- Plot 2: proportion_10_baseline vs. proportion_10_SA2 ---
plot2 <- ggplot(data = lsoa_summary, aes(x = proportion_10_baseline, y = proportion_10_SA2)) +
  geom_point(alpha = 0.9, color = "#DDB0D9") +
  geom_point(size = 3, alpha = 0.9, shape = 21, stroke = 0.25, color = "black") +
  geom_abline(intercept = 0, slope = 1, linetype = "dashed", color = "black") +
  scale_x_continuous(breaks = seq(0, 100, 10), limits = c(0, 100)) +
  scale_y_continuous(breaks = seq(0, 100, 10), limits = c(0, 100)) +
  labs(
    title = "LSOA-level 10% fuel poverty: Baseline vs. SA2 conditions",
    x = "Baseline fuel poverty (%)",
    y = "SA2 fuel poverty (%)"
  ) +
  theme_bw() +
  theme(axis.text.x = element_text(size = 17),
        axis.text.y = element_text(size = 17),
        legend.text = element_text(size = 17),
        legend.title = element_text(size = 17),
        plot.title = element_text(size = 18),
        axis.title.x = element_text(size = 17),
        axis.title.y = element_text(size = 17)) +
  coord_fixed()

# print(plot2)

grid_10 <- plot_grid(plot1, plot2, nrow = 1, align = "g")
print(grid_10)

# --- Plot 3: proportion_lihc_baseline vs. proportion_lihc_SA1 ---
plot3 <- ggplot(data = lsoa_summary, aes(x = proportion_lihc_baseline, y = proportion_lihc_SA1)) +
  geom_point(alpha = 0.9, color = "#ACE4E4") +
  geom_point(size = 3, alpha = 0.9, shape = 21, stroke = 0.25, color = "black") +
  geom_abline(intercept = 0, slope = 1, linetype = "dashed", color = "black") +
  scale_x_continuous(breaks = seq(0, 90, 10), limits = c(0, 90)) +
  scale_y_continuous(breaks = seq(0, 90, 10), limits = c(0, 90)) +
  labs(
    title = "LSOA-level LIHC fuel poverty: Baseline vs. SA1 conditions",
    x = "Baseline fuel poverty (%)",
    y = "SA1 fuel poverty (%)"
  ) +
  theme_bw() +
  theme(axis.text.x = element_text(size = 17),
        axis.text.y = element_text(size = 17),
        legend.text = element_text(size = 17),
        legend.title = element_text(size = 17),
        plot.title = element_text(size = 18),
        axis.title.x = element_text(size = 17),
        axis.title.y = element_text(size = 17)) +
  coord_fixed()

# print(plot3)

# --- Plot 4: proportion_lihc_baseline vs. proportion_lihc_SA2 ---
plot4 <- ggplot(data = lsoa_summary, aes(x = proportion_lihc_baseline, y = proportion_lihc_SA2)) +
  geom_point(alpha = 0.9, color = "#ACE4E4") +
  geom_point(size = 3, alpha = 0.9, shape = 21, stroke = 0.25, color = "black") +
  geom_abline(intercept = 0, slope = 1, linetype = "dashed", color = "black") +
  scale_x_continuous(breaks = seq(0, 90, 10), limits = c(0, 90)) +
  scale_y_continuous(breaks = seq(0, 90, 10), limits = c(0, 90)) +
  labs(
    title = "LSOA-level LIHC fuel poverty: Baseline vs. SA2 conditions",
    x = "Baseline fuel poverty (%)",
    y = "SA2 fuel poverty (%)"
  ) +
  theme_bw() +
  theme(axis.text.x = element_text(size = 17),
        axis.text.y = element_text(size = 17),
        legend.text = element_text(size = 17),
        legend.title = element_text(size = 17),
        plot.title = element_text(size = 18),
        axis.title.x = element_text(size = 17),
        axis.title.y = element_text(size = 17)) +
  coord_fixed()

# print(plot4)

grid_lihc <- plot_grid(plot3, plot4, nrow = 1, align = "g")
print(grid_lihc)

# --- Plot 5: proportion_lilee_baseline vs. proportion_lilee_SA1 ---
plot5 <- ggplot(data = lsoa_summary, aes(x = proportion_lilee_baseline, y = proportion_lilee_SA1)) +
  geom_point(alpha = 0.9, color = "#5A97BC") +
  geom_point(size = 3, alpha = 0.9, shape = 21, stroke = 0.25, color = "black") +
  geom_abline(intercept = 0, slope = 1, linetype = "dashed", color = "black") +
  scale_x_continuous(breaks = seq(0, 70, 10), limits = c(0, 70)) +
  scale_y_continuous(breaks = seq(0, 70, 10), limits = c(0, 70)) +
  labs(
    title = "LSOA-level LILEE fuel poverty: Baseline vs. SA1 conditions",
    x = "Baseline fuel poverty (%)",
    y = "SA1 fuel poverty (%)"
  ) +
  theme_bw() +
  theme(axis.text.x = element_text(size = 17),
        axis.text.y = element_text(size = 17),
        legend.text = element_text(size = 17),
        legend.title = element_text(size = 17),
        plot.title = element_text(size = 18),
        axis.title.x = element_text(size = 17),
        axis.title.y = element_text(size = 17)) +
  coord_fixed()


# print(plot5)

# --- Plot 6: proportion_lilee_baseline vs. proportion_lilee_SA2 ---
plot6 <- ggplot(data = lsoa_summary, aes(x = proportion_lilee_baseline, y = proportion_lilee_SA2)) +
  geom_point(alpha = 0.9, color = "#5A97BC") +
  geom_point(size = 3, alpha = 0.9, shape = 21, stroke = 0.25, color = "black") +
  geom_abline(intercept = 0, slope = 1, linetype = "dashed", color = "black") +
  scale_x_continuous(breaks = seq(0, 70, 10), limits = c(0, 70)) +
  scale_y_continuous(breaks = seq(0, 70, 10), limits = c(0, 70)) +
  labs(
    title = "LSOA-level LILEE fuel poverty: Baseline vs. SA2 conditions",
    x = "Baseline fuel poverty (%)",
    y = "SA2 fuel poverty (%)"
  ) +
  theme_bw() +
  theme(axis.text.x = element_text(size = 15),
        axis.text.y = element_text(size = 15),
        legend.text = element_text(size = 15),
        legend.title = element_text(size = 15),
        plot.title = element_text(size = 16),
        axis.title.x = element_text(size = 15),
        axis.title.y = element_text(size = 15)) +
  coord_fixed()

# print(plot6)

grid_lilee <- plot_grid(plot5, plot6, nrow = 1, align = "g")
print(grid_lilee)

# --- Plot 7: proportion_mis_plus_10_baseline vs. proportion_mis_plus_10_SA1 ---
plot7 <- ggplot(data = lsoa_summary, aes(x = proportion_mis_plus_10_baseline, y = proportion_mis_plus_10_SA1)) +
  geom_point(alpha = 0.9, color = "#BB64B1") +
  geom_point(size = 3, alpha = 0.9, shape = 21, stroke = 0.25, color = "black") +
  geom_abline(intercept = 0, slope = 1, linetype = "dashed", color = "black") +
  scale_x_continuous(breaks = seq(0, 100, 10), limits = c(0, 100)) +
  scale_y_continuous(breaks = seq(0, 100, 10), limits = c(0, 100)) +
  labs(
    title = "LSOA-level 10% + MIS fuel poverty: Baseline vs. SA1 conditions",
    x = "Baseline fuel poverty (%)",
    y = "SA1 fuel poverty (%)"
  ) +
  theme_bw() +
  theme(axis.text.x = element_text(size = 17),
        axis.text.y = element_text(size = 17),
        legend.text = element_text(size = 17),
        legend.title = element_text(size = 17),
        plot.title = element_text(size = 18),
        axis.title.x = element_text(size = 17),
        axis.title.y = element_text(size = 17)) +
  coord_fixed()


# print(plot7)

# --- Plot 8: proportion_mis_plus_10_baseline vs. proportion_mis_plus_10_SA2 ---
plot8 <- ggplot(data = lsoa_summary, aes(x = proportion_mis_plus_10_baseline, y = proportion_mis_plus_10_SA2)) +
  geom_point(alpha = 0.9, color = "#BB64B1") +
  geom_point(size = 3, alpha = 0.9, shape = 21, stroke = 0.25, color = "black") +
  geom_abline(intercept = 0, slope = 1, linetype = "dashed", color = "black") +
  scale_x_continuous(breaks = seq(0, 100, 10), limits = c(0, 100)) +
  scale_y_continuous(breaks = seq(0, 100, 10), limits = c(0, 100)) +
  labs(
    title = "LSOA-level 10% + MIS fuel poverty: Baseline vs. SA2 conditions",
    x = "Baseline fuel poverty (%)",
    y = "SA2 fuel poverty (%)"
  ) +
  theme_bw() +
  theme(axis.text.x = element_text(size = 17),
        axis.text.y = element_text(size = 17),
        legend.text = element_text(size = 17),
        legend.title = element_text(size = 17),
        plot.title = element_text(size = 18),
        axis.title.x = element_text(size = 17),
        axis.title.y = element_text(size = 17)) +
  coord_fixed()

# print(plot8)

grid_mis_10 <- plot_grid(plot7, plot8, nrow = 1, align = "g")
print(grid_mis_10)


```


## Calculate LSOA-level 10% fuel poverty quotient for baseline, SA1 and SA2
```{r}

# This is a measure of energy equality, equivalent to aggregate fuel poverty gap / fuel poverty buffer per area

synth_fp_quotient <- synth_fp

# BASELINE FP QUOTIENT
# First calculate aggregate 10% gap and buffer per LSOA
lsoa_fp_quotient <- synth_fp_quotient %>%
  group_by(lsoa_2011) %>%
  summarise(
    aggregate_10_gap_baseline = sum(`X10._fuel_poverty_gap_and_buffer`[`X10._fuel_poverty_gap_and_buffer` > 0], na.rm = TRUE),
    aggregate_10_buffer_baseline = sum(`X10._fuel_poverty_gap_and_buffer`[`X10._fuel_poverty_gap_and_buffer` < 0], na.rm = TRUE)
  ) %>%
  ungroup()

# Now calculate 10% fuel poverty quotient
lsoa_fp_quotient <- lsoa_fp_quotient %>%
  mutate(
    `quotient_10_baseline` = ifelse(
      aggregate_10_buffer_baseline == 0,
      NA, # Assign NA if the denominator is zero
      aggregate_10_gap_baseline / abs(aggregate_10_buffer_baseline) # Note: 'abs' ensures that all buffer values are treated as positive values for quotient calculation
    )
  )





# SENSITIVITY ANALYSIS 1 FP QUOTIENT

# 10% gap (positive values) and buffer (negative values)
synth_fp_quotient <- synth_fp_quotient %>%
  mutate(`10_fuel_poverty_gap_and_buffer_SA1` = energy_costs_SA1 - (0.10 * income))

fp_quotient_SA1 <- synth_fp_quotient %>%
  group_by(lsoa_2011) %>%
  summarise(
    aggregate_10_gap_SA1 = sum(`10_fuel_poverty_gap_and_buffer_SA1`[`10_fuel_poverty_gap_and_buffer_SA1` > 0], na.rm = TRUE),
    aggregate_10_buffer_SA1 = sum(`10_fuel_poverty_gap_and_buffer_SA1`[`10_fuel_poverty_gap_and_buffer_SA1` < 0], na.rm = TRUE)
  ) %>%
  ungroup()

# Now calculate 10% fuel poverty quotient
fp_quotient_SA1 <- fp_quotient_SA1 %>%
  mutate(
    `quotient_10_SA1` = ifelse(
      aggregate_10_buffer_SA1 == 0,
      NA, # Assign NA if the denominator is zero
      aggregate_10_gap_SA1 / abs(aggregate_10_buffer_SA1) # Note: 'abs' ensures that all buffer values are treated as positive values for quotient calculation
    )
  )





# SENSITIVITY ANALYSIS 2 FP QUOTIENT

# 10% gap (positive values) and buffer (negative values)
synth_fp_quotient <- synth_fp_quotient %>%
  mutate(`10_fuel_poverty_gap_and_buffer_SA2` = energy_costs_SA2 - (0.10 * income_SA2))

fp_quotient_SA2 <- synth_fp_quotient %>%
  group_by(lsoa_2011) %>%
  summarise(
    aggregate_10_gap_SA2 = sum(`10_fuel_poverty_gap_and_buffer_SA2`[`10_fuel_poverty_gap_and_buffer_SA2` > 0], na.rm = TRUE),
    aggregate_10_buffer_SA2 = sum(`10_fuel_poverty_gap_and_buffer_SA2`[`10_fuel_poverty_gap_and_buffer_SA2` < 0], na.rm = TRUE)
  ) %>%
  ungroup()

# Now calculate 10% fuel poverty quotient
fp_quotient_SA2 <- fp_quotient_SA2 %>%
  mutate(
    `quotient_10_SA2` = ifelse(
      aggregate_10_buffer_SA2 == 0,
      NA, # Assign NA if the denominator is zero
      aggregate_10_gap_SA2 / abs(aggregate_10_buffer_SA2) # Note: 'abs' ensures that all buffer values are treated as positive values for quotient calculation
    )
  )



# Left join to form one LSOA quotient summary
fp_quotient_summary <- lsoa_fp_quotient %>%
  left_join(fp_quotient_SA1, by = "lsoa_2011") %>%
  left_join(fp_quotient_SA2, by = "lsoa_2011")


# Too cumbersome to display results for LSOAs; therefore, results are demonstrated for the previously used LSOA subset (as shown below)
lsoa_values_to_keep <- c("E01013979", "E01013983", "E01013931", "E01013965", 
                          "E01013977", "E01013893", "E01013897", "E01013856", 
                          "E01013934", "E01013984", "E01033398")

fp_quotient_subset <- fp_quotient_summary %>%
  filter(lsoa_2011 %in% lsoa_values_to_keep)


# write.csv(x=fp_quotient_subset, file="10_fuel_poverty_quotient_subset_summary.csv")


# Nottingham total FP quotient summaries

# Baseline
Notts_base_quo <- synth_fp_quotient %>%
  summarise(
    aggregate_10_gap_baseline = sum(`X10._fuel_poverty_gap_and_buffer`[`X10._fuel_poverty_gap_and_buffer` > 0], na.rm = TRUE),
    aggregate_10_buffer_baseline = sum(`X10._fuel_poverty_gap_and_buffer`[`X10._fuel_poverty_gap_and_buffer` < 0], na.rm = TRUE)
  )


# SA1
Notts_SA1_quo <- synth_fp_quotient %>%
  summarise(
    aggregate_10_gap_SA1 = sum(`10_fuel_poverty_gap_and_buffer_SA1`[`10_fuel_poverty_gap_and_buffer_SA1` > 0], na.rm = TRUE),
    aggregate_10_buffer_SA1 = sum(`10_fuel_poverty_gap_and_buffer_SA1`[`10_fuel_poverty_gap_and_buffer_SA1` < 0], na.rm = TRUE)
  )


#SA2
Notts_SA2_quo <- synth_fp_quotient %>%
    summarise(
    aggregate_10_gap_SA2 = sum(`10_fuel_poverty_gap_and_buffer_SA2`[`10_fuel_poverty_gap_and_buffer_SA2` > 0], na.rm = TRUE),
    aggregate_10_buffer_SA2 = sum(`10_fuel_poverty_gap_and_buffer_SA2`[`10_fuel_poverty_gap_and_buffer_SA2` < 0], na.rm = TRUE)
  )


```


